Density Matrix Susceptibility

### 2018-05-22 papers/1_dmatpol/paper-r1.tex
RMSE indicator coded and working well

In order to validate all the models discussed in this Contribution,
the electric field\hyp{}induced interaction energy, 
the dipole moment and the quadrupole moment were analysed in detail.
The electric field\hyp{}induced interaction energy at the HF level, $\delta E$,
was computed from the following exact formula:
%
\begin{multline}\label{e:dE}
 \delta E[{\bf F}] = 2\Tr{\left\{ {\bf D}^{(0)} \cdot {\bf V}[{\bf F}] \right\}}
                   + \Tr{\left\{ \left( {\bf D}^{(0)} + \delta {\bf D} \right) \cdot
                                \left( \delta {\bf J} + \delta {\bf K}\right) \right\}} \\
                   + \Tr{\left\{ \delta {\bf D} \cdot
                                \left( {\bf H}^{(0)} + 2 {\bf V}[{\bf F}] + {\bf f}^{(0)} \right) \right\}}
                   + E_{\rm nuc}({\bf F}) \;.
\end{multline}
%
In Eq.~\eqref{e:dE}, ${\bf f}^{(0)}$ and ${\bf H}^{(0)}$ are the Fock matrix and 
the one\hyp{}electron Hamiltonian matrix of an unperturbed HF ground state,
${\bf V}[{\bf F}]$ is the one\hyp{}electron external electrostatic potential matrix whereas $\delta {\bf J}$
and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
the polarization of the electron density distribution, respectively. $E_{\rm nuc}({\bf F})$ denotes the interaction energy
due to the nuclei.
The electric field\hyp{}induced dipole and quadrupole moments were evaluated from the following expressions:
%
\begin{subequations}\label{e:dmult}
  \begin{align}
   \delta \upmu_{u}   &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{u}{\beta}  \;,\\
   \delta \Theta_{uw} &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{uw}{\beta} \;. 
  \end{align}
\end{subequations}
%
To compute the interaction energy and the induced multipole moments, either exact or approximated forms
of the difference density matrix from Eq.~\eqref{e:final-model.General}
were used for a particular electric field distribution. The statistical evaluation of
errors with respect to the benchmark (full HF calculations) were based on 
the following statistical quality descriptors:
%
\begin{enumerate}
 \item Root Mean Square Energy Error (RMSE)
   \begin{equation}
     {\rm RMSE} = \sum_N \sqrt{ \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 }
   \end{equation}
 \item Root Mean Square Dipole Error (RMSD)
   \begin{equation}
     {\rm RMSD} = \sum_N \sqrt{ \lvert \delta {\BM\upmu}^{(N)} - \overline{\delta {\BM\upmu}^{(N)}}\rvert^2 }
   \end{equation}
 \item Root Mean Square Quadrupole Error (RMSQ)
   \begin{equation}
     {\rm RMSQ} = \sum_N \sqrt{ \lvert \delta {\BM\Theta}^{(N)} - \overline{\delta {\BM\Theta}^{(N)}}\rvert^2 }
   \end{equation}
\end{enumerate}
%
The first describes the ability of the model to reproduce the exact energetics of the system, 
wherease the latter two contain crucial information on how well the deformation of the one\hyp{}electron
density is described.

>>>>

In order to validate all the models discussed in this Contribution,
the electric field\hyp{}induced interaction energy, 
the dipole moment and the quadrupole moment were analysed in detail.
The electric field\hyp{}induced interaction energy at the HF level, $\delta E$,
was computed from the following exact formula:
%
\begin{multline}\label{e:dE}
 \delta E[{\bf F}] = 
                     \Tr{\left\{ \delta {\bf D} \cdot
                                \left( {\bf H}^{(0)} + {\bf f}^{(0)} \right) \right\}}
                   + \Tr{\left\{ \left( {\bf D}^{(0)} + \delta {\bf D} \right) \cdot
                                \left( 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} 
                   + E_{\rm nuc}({\bf F}) \;.
\end{multline}
%
In Eq.~\eqref{e:dE}, ${\bf f}^{(0)}$ and ${\bf H}^{(0)}$ are the Fock matrix and 
the one\hyp{}electron Hamiltonian matrix of an unperturbed HF ground state,
${\bf V}$ is the one\hyp{}electron external electrostatic potential matrix whereas $\delta {\bf J}$
and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
the polarization of the electron density distribution, respectively. $E_{\rm nuc}({\bf F})$ denotes the interaction energy
due to the nuclei.
The electric field\hyp{}induced dipole and quadrupole moments were evaluated from the following expressions:
%
\begin{subequations}\label{e:dmult}
  \begin{align}
   \delta \upmu_{u}   &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{u}{\beta}  \;,\\
   \delta \Theta_{uw} &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{uw}{\beta} \;. 
  \end{align}
\end{subequations}
%
To compute the interaction energy and the induced multipole moments, either exact or approximated forms
of the difference density matrix from Eq.~\eqref{e:final-model.General}
were used for a particular electric field distribution. The statistical evaluation of
errors with respect to the benchmark (full HF calculations) were based on 
the following statistical quality descriptors:
%
\begin{enumerate}
 \item Root Mean Square Energy Error (RMSE)
   \begin{equation}
     {\rm RMSE} = \sqrt{ \frac{1}{N_{\rm max}} \sum_N \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 }
   \end{equation}
 \item Root Mean Square Dipole Error (RMSD)
   \begin{equation}
     {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta {\BM\upmu}^{(N)} - \overline{\delta {\BM\upmu}^{(N)}}\rvert^2 }
   \end{equation}
 \item Root Mean Square Quadrupole Error (RMSQ)
   \begin{equation}
     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N \lvert \delta {\BM\Theta}^{(N)} - \overline{\delta {\BM\Theta}^{(N)}}\rvert^2 }
   \end{equation}
\end{enumerate}
%
The first describes the ability of the model to reproduce the exact energetics of the system, 
wherease the latter two contain crucial information on how well the deformation of the one\hyp{}electron
density is described.
